---
title: "DREADD PFC Silencing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup libraries and paths

```{r, warning=FALSE, message=FALSE}
# load libraries you'll need
library(easypackages)
libraries("nlme","ggplot2","here","matlabr","reshape2")

# define paths
codepath = here("code")
phenopath  = here("data","dreadd")
```

## Run MATLAB code to estimate H

```{r, warning=FALSE, message=FALSE}
RUNMATLAB = FALSE
if(RUNMATLAB){
  code2run = sprintf("cd %s; B_invivo_2_DREADDpfc_silencing;",codepath)
  res_tmp = run_matlab_code(code2run)
}
```


## Setup data for analysis

```{r, warning=FALSE, message=FALSE}
# initial information
window_size = 512
nwindows = 3389
vol_inject_start = 900
vol_treatment_start = 1800

# read in sliding-window analysis data
data_win = read.csv(file.path(phenopath,"pheno_data+Hwin_dreaddsilencing.csv"))

# melt the data frame into long format
data_win_long = melt(data_win, 
                     id.vars = c("filename","condition","scan_day","H"))
colnames(data_win_long)[ncol(data_win_long)] = "Hwin"

# make a time column
data_win_long$time = NA
for (i in 1:nwindows){
  mask = data_win_long$variable==sprintf("window_%04d",i)
  data_win_long[mask,"time"] = i
}

# make column indicating treatment phase
data_win_long$treatment_phase = NA
data_win_long$treatment_phase[data_win_long$time<=(vol_inject_start-window_size)-1] = "Baseline"
data_win_long$treatment_phase[data_win_long$time>=(vol_inject_start-window_size) & data_win_long$time<=(vol_treatment_start-window_size)-1] = "Transition"
data_win_long$treatment_phase[data_win_long$time>(vol_treatment_start-window_size)] = "Treatment"
data_win_long$treatment_phase = factor(data_win_long$treatment_phase)
```

## Main linear mixed effect model on sliding window H data

```{r, warning=FALSE, message=FALSE}
# pick data to use in model
df2use = data_win_long

# formulas for fixed and random effects
fx_form = as.formula("Hwin ~ scan_day + condition*time*treatment_phase")
rx_form = as.formula("~ time|filename + treatment_phase|filename")

# other things to include in the call to lme
ctrl = lmeControl(opt='optim', msMaxIter = 500)
ml_method = "REML"

# run the lme model
mod2use = eval(substitute(lme(fixed = fx_form, 
                               random = rx_form, 
                               data = df2use, 
                               na.action = na.omit, 
                               control=ctrl, 
                               method = ml_method),
                           list(fx_form = fx_form, rx_form = rx_form)))

# compute ANOVA stats from the model
anova(mod2use)
```

## Decomposing the Time x DREADD x Treatment Phase interaction

### Baseline

```{r, warning=FALSE, message=FALSE}
# grab data from Baseline period
df2use = subset(data_win_long, data_win_long$treatment_phase=="Baseline")

# formulas
fx_form = as.formula("Hwin ~ scan_day + condition*time")
rx_form = as.formula("~ time|filename")

# run model
mod2use = eval(substitute(lme(fixed = fx_form, 
                               random = rx_form, 
                               data = df2use, 
                               na.action = na.omit, 
                               control=ctrl, 
                               method = ml_method),
                           list(fx_form = fx_form, rx_form = rx_form)))

# ANOVA stats on the model
anova(mod2use)

# compute descriptive stats - mean and SD
descrip_res = data.frame(matrix(nrow = 2, ncol = 2))
rownames(descrip_res) = c("DREADD","SHAM")
colnames(descrip_res) = c("Mean","SD")
descrip_res["DREADD","Mean"] = mean(df2use[df2use$condition=="DREADD","Hwin"], na.rm = TRUE)
descrip_res["SHAM","Mean"] = mean(df2use[df2use$condition=="SHAM","Hwin"], na.rm = TRUE)
descrip_res["DREADD","SD"] = sd(df2use[df2use$condition=="DREADD","Hwin"], na.rm = TRUE)
descrip_res["SHAM","SD"] = sd(df2use[df2use$condition=="SHAM","Hwin"], na.rm = TRUE)
descrip_res
```

### Transition

```{r, warning=FALSE, message=FALSE}
# grab data for Transition period
df2use = subset(data_win_long, data_win_long$treatment_phase=="Transition")

# formulas
fx_form = as.formula("Hwin ~  scan_day + condition*time")
rx_form = as.formula("~ time|filename")

# run model
mod2use = eval(substitute(lme(fixed = fx_form, 
                               random = rx_form, 
                               data = df2use, 
                               na.action = na.omit, 
                               control=ctrl, 
                               method = ml_method),
                          list(fx_form = fx_form, rx_form = rx_form)))

# ANOVA stats
anova(mod2use)

# compute descriptives
descrip_res = data.frame(matrix(nrow = 2, ncol = 2))
rownames(descrip_res) = c("DREADD","SHAM")
colnames(descrip_res) = c("Mean","SD")
descrip_res["DREADD","Mean"] = mean(df2use[df2use$condition=="DREADD","Hwin"], na.rm = TRUE)
descrip_res["SHAM","Mean"] = mean(df2use[df2use$condition=="SHAM","Hwin"], na.rm = TRUE)
descrip_res["DREADD","SD"] = sd(df2use[df2use$condition=="DREADD","Hwin"], na.rm = TRUE)
descrip_res["SHAM","SD"] = sd(df2use[df2use$condition=="SHAM","Hwin"], na.rm = TRUE)
descrip_res
```

### Treatment

```{r, warning=FALSE, message=FALSE}
# grab data from Treatment period
df2use = subset(data_win_long, data_win_long$treatment_phase=="Treatment")

# formulas
fx_form = as.formula("Hwin ~  scan_day + condition*time")
rx_form = as.formula("~ time|filename")

# run model
mod2use = eval(substitute(lme(fixed = fx_form, 
                               random = rx_form, 
                               data = df2use, 
                               na.action = na.omit, 
                               control=ctrl, 
                               method = ml_method),
                           list(fx_form = fx_form, rx_form = rx_form)))

# ANOVA stats
anova(mod2use)

# compute descriptives
descrip_res = data.frame(matrix(nrow = 2, ncol = 2))
rownames(descrip_res) = c("DREADD","SHAM")
colnames(descrip_res) = c("Mean","SD")
descrip_res["DREADD","Mean"] = mean(df2use[df2use$condition=="DREADD","Hwin"], na.rm = TRUE)
descrip_res["SHAM","Mean"] = mean(df2use[df2use$condition=="SHAM","Hwin"], na.rm = TRUE)
descrip_res["DREADD","SD"] = sd(df2use[df2use$condition=="DREADD","Hwin"], na.rm = TRUE)
descrip_res["SHAM","SD"] = sd(df2use[df2use$condition=="SHAM","Hwin"], na.rm = TRUE)
descrip_res
```

## Plot

```{r, warning=FALSE, message=FALSE}
# plotting using GAM smoother for both individual and group trajectories
yLims = c(1.4,1.8)
p = ggplot(data = data_win_long, 
           aes(x = time,
               y = Hwin,
               group = filename)) +
            facet_grid(. ~ condition)

# add smoothed individual lines for each mouse
p = p + geom_smooth(se = FALSE ,colour='gray75',alpha=0.1)

# add group-level fit line
p = p + geom_smooth(se=TRUE, aes(group = interaction(condition,treatment_phase), colour = treatment_phase), size=4)

# add other stuff to the plot
p = p + geom_vline(xintercept = (vol_inject_start-window_size)) +
  geom_vline(xintercept = vol_treatment_start-window_size) +
  ylab("Hurst Exponent (H)") +
  xlab("Time Window") +
  scale_y_continuous(limits = yLims)

# save figure
# ggsave(filename = here("figures","DREADDsilencing_pfc_sliding_window_fig_GAMsmooth.pdf"))
p
```



# fALFF

## Setup data for analysis

```{r, warning=FALSE, message=FALSE}
# initial information
window_size = 512
nwindows = 3389
vol_inject_start = 900
vol_treatment_start = 1800

# read in sliding-window analysis data
data_win = read.csv(file.path(phenopath,"pheno_data+fALFFwin_dreaddsilencing.csv"))

# melt the data frame into long format
data_win_long = melt(data_win, 
                     id.vars = c("filename","condition","scan_day"))
colnames(data_win_long)[ncol(data_win_long)] = "fALFFwin"

# make a time column
data_win_long$time = NA
for (i in 1:nwindows){
  mask = data_win_long$variable==sprintf("window_%04d",i)
  data_win_long[mask,"time"] = i
}

# make column indicating treatment phase
data_win_long$treatment_phase = NA
data_win_long$treatment_phase[data_win_long$time<=(vol_inject_start-window_size)-1] = "Baseline"
data_win_long$treatment_phase[data_win_long$time>=(vol_inject_start-window_size) & data_win_long$time<=(vol_treatment_start-window_size)-1] = "Transition"
data_win_long$treatment_phase[data_win_long$time>(vol_treatment_start-window_size)] = "Treatment"
data_win_long$treatment_phase = factor(data_win_long$treatment_phase)
```

## Main linear mixed effect model on sliding window H data

```{r, warning=FALSE, message=FALSE}
# pick data to use in model
df2use = data_win_long

# formulas for fixed and random effects
fx_form = as.formula("fALFFwin ~ scan_day + condition*time*treatment_phase")
rx_form = as.formula("~ time|filename + treatment_phase|filename")

# other things to include in the call to lme
ctrl = lmeControl(opt='optim', msMaxIter = 500)
ml_method = "REML"

# run the lme model
mod2use = eval(substitute(lme(fixed = fx_form, 
                               random = rx_form, 
                               data = df2use, 
                               na.action = na.omit, 
                               control=ctrl, 
                               method = ml_method),
                           list(fx_form = fx_form, rx_form = rx_form)))

# compute ANOVA stats from the model
anova(mod2use)
```

## Decomposing the Time x DREADD x Treatment Phase interaction

### Baseline

```{r, warning=FALSE, message=FALSE}
# grab data from Baseline period
df2use = subset(data_win_long, data_win_long$treatment_phase=="Baseline")

# formulas
fx_form = as.formula("fALFFwin ~ scan_day + condition*time")
rx_form = as.formula("~ time|filename")

# run model
mod2use = eval(substitute(lme(fixed = fx_form, 
                               random = rx_form, 
                               data = df2use, 
                               na.action = na.omit, 
                               control=ctrl, 
                               method = ml_method),
                           list(fx_form = fx_form, rx_form = rx_form)))

# ANOVA stats on the model
anova(mod2use)

# compute descriptive stats - mean and SD
descrip_res = data.frame(matrix(nrow = 2, ncol = 2))
rownames(descrip_res) = c("DREADD","SHAM")
colnames(descrip_res) = c("Mean","SD")
descrip_res["DREADD","Mean"] = mean(df2use[df2use$condition=="DREADD","fALFFwin"], na.rm = TRUE)
descrip_res["SHAM","Mean"] = mean(df2use[df2use$condition=="SHAM","fALFFwin"], na.rm = TRUE)
descrip_res["DREADD","SD"] = sd(df2use[df2use$condition=="DREADD","fALFFwin"], na.rm = TRUE)
descrip_res["SHAM","SD"] = sd(df2use[df2use$condition=="SHAM","fALFFwin"], na.rm = TRUE)
descrip_res
```

### Transition

```{r, warning=FALSE, message=FALSE}
# grab data for Transition period
df2use = subset(data_win_long, data_win_long$treatment_phase=="Transition")

# formulas
fx_form = as.formula("fALFFwin ~  scan_day + condition*time")
rx_form = as.formula("~ time|filename")

# run model
mod2use = eval(substitute(lme(fixed = fx_form, 
                               random = rx_form, 
                               data = df2use, 
                               na.action = na.omit, 
                               control=ctrl, 
                               method = ml_method),
                          list(fx_form = fx_form, rx_form = rx_form)))

# ANOVA stats
anova(mod2use)

# compute descriptives
descrip_res = data.frame(matrix(nrow = 2, ncol = 2))
rownames(descrip_res) = c("DREADD","SHAM")
colnames(descrip_res) = c("Mean","SD")
descrip_res["DREADD","Mean"] = mean(df2use[df2use$condition=="DREADD","fALFFwin"], na.rm = TRUE)
descrip_res["SHAM","Mean"] = mean(df2use[df2use$condition=="SHAM","fALFFwin"], na.rm = TRUE)
descrip_res["DREADD","SD"] = sd(df2use[df2use$condition=="DREADD","fALFFwin"], na.rm = TRUE)
descrip_res["SHAM","SD"] = sd(df2use[df2use$condition=="SHAM","fALFFwin"], na.rm = TRUE)
descrip_res
```

### Treatment

```{r, warning=FALSE, message=FALSE}
# grab data from Treatment period
df2use = subset(data_win_long, data_win_long$treatment_phase=="Treatment")

# formulas
fx_form = as.formula("fALFFwin ~  scan_day + condition*time")
rx_form = as.formula("~ time|filename")

# run model
mod2use = eval(substitute(lme(fixed = fx_form, 
                               random = rx_form, 
                               data = df2use, 
                               na.action = na.omit, 
                               control=ctrl, 
                               method = ml_method),
                           list(fx_form = fx_form, rx_form = rx_form)))

# ANOVA stats
anova(mod2use)

# compute descriptives
descrip_res = data.frame(matrix(nrow = 2, ncol = 2))
rownames(descrip_res) = c("DREADD","SHAM")
colnames(descrip_res) = c("Mean","SD")
descrip_res["DREADD","Mean"] = mean(df2use[df2use$condition=="DREADD","fALFFwin"], na.rm = TRUE)
descrip_res["SHAM","Mean"] = mean(df2use[df2use$condition=="SHAM","fALFFwin"], na.rm = TRUE)
descrip_res["DREADD","SD"] = sd(df2use[df2use$condition=="DREADD","fALFFwin"], na.rm = TRUE)
descrip_res["SHAM","SD"] = sd(df2use[df2use$condition=="SHAM","fALFFwin"], na.rm = TRUE)
descrip_res
```

## Plot

```{r, warning=FALSE, message=FALSE}
# plotting using GAM smoother for both individual and group trajectories
yLims = c(0,1)
p = ggplot(data = data_win_long, 
           aes(x = time,
               y = fALFFwin,
               group = filename)) +
            facet_grid(. ~ condition)

# add smoothed individual lines for each mouse
p = p + geom_smooth(se = FALSE ,colour='gray75',alpha=0.1)

# add group-level fit line
p = p + geom_smooth(se=TRUE, aes(group = interaction(condition,treatment_phase), colour = treatment_phase), size=5)

# add other stuff to the plot
p = p + geom_vline(xintercept = (vol_inject_start-window_size)) +
  geom_vline(xintercept = vol_treatment_start-window_size) +
  ylab("fALFF") +
  xlab("Time Window") +
  scale_y_continuous(limits = yLims)

# save figure
# ggsave(filename = here("figures","DREADDsilencing_pfc_sliding_window_fig_GAMsmooth_fALFF.pdf"))
p
```
