---
title: "Hurst PLS Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r, warning=FALSE, message=FALSE}
# load libraries you'll need
library(easypackages)
libraries("readxl","ggplot2","MASS","psych","here","heplots","ggseg","patchwork","knitr","matlabr")
source(here("github_repo","code","cohens_d.R"))

# FDR threshold to use
fdr_thresh = 0.05

# define paths
codepath = here("github_repo","code")
phenopath  = here("pheno")
datapath = here("data")
resultpath = here("pls_results")

# load HCP annotations
hcp_annot = read.delim(file.path(codepath, "GlasserHCP_annot.txt"))
```

Run MATLAB code to do PLS analysis

```{r, warning=FALSE, message=FALSE}
RUNMATLAB = FALSE
if(RUNMATLAB){
  code2run = sprintf("cd %s; C_3a_AIMS_Hurst_PLS;",codepath)
  res_tmp = run_matlab_code(code2run)
}
```

PLS analysis where the Y matrix is a contrast matrix that includes both main effects of sex and diagnosis, as well as the sex*diagnosis interaction.

```{r, warning=FALSE, message=FALSE}
data2plot = read.csv(file.path(resultpath,"pls_ALL_H_bootCI4plotting_LV1_ci95.csv"))

p = ggplot(data = data2plot, aes(x = Effect, y = corr, colour = Effect))
p = p + geom_pointrange(aes(ymin = lo_lim, ymax = up_lim))
p = p + geom_linerange(data = data2plot, aes(ymin = lo_lim, ymax = up_lim))
p = p + ylab("Correlation")
p
```

PLS analysis on the specific sex*diagnosis contrast 

```{r, warning=FALSE, message=FALSE}
data2plot = read.csv(file.path(resultpath,"pls_SEXbyDX_H_brainBSR4plotting.csv"))
data2plot =  merge(data2plot,hcp_annot,by="parcel_index")

data2plot$BSR_LV1_thresh = data2plot$BSR_LV1
mask = abs(data2plot$BSR_LV1_thresh)>abs(quantile(data2plot$BSR_LV1, c(0.8)))
data2plot$BSR_LV1_thresh[!mask] = NA

data2plot = data2plot[-c(131,172),]

# show effect size throughout the entire brain
g1 = ggseg(data = data2plot, mapping=aes(fill = BSR_LV1), position = "stacked", atlas = "glasser") + theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="BSR")) +
  scale_fill_gradient2(midpoint=0, low="blue", mid="white", high="red", limits=c(-25, 55), breaks = round(seq(from=-25,to=55, by=25),2)) + guides(fill = guide_colourbar(nbin = 100))
g1

# show top 20% of brain regions ranked by BSR
g2 = ggseg(data = data2plot, mapping=aes(fill = BSR_LV1_thresh), position = "stacked", atlas = "glasser") + theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="BSR")) +
  scale_fill_gradient2(midpoint=0, low="blue", mid="white", high="red", limits=c(-25, 55), breaks = round(seq(from=-25,to=55, by=25),2)) + guides(fill = guide_colourbar(nbin = 100))
g2
```

Plot a histogram of the correlation between H and the sex*diagnosis interaction effect across all bootstrap 10,000 resamples. The actual correlation in the real data is shown with the vertical line. This analysis will show us descriptively the correlation between H and the interaction effect, as well as showing that in bootstrap analysis that this correlation is indeed non-zero, as the previous analysis already showed us. 

```{r, warning=FALSE, message=FALSE}
data2plot = read.csv(file.path(resultpath,"pls_SEXbyDX_H_bootres4plotting_LV1_ci95.csv"))

p = ggplot(data = data2plot, aes(x = boot_res)) + geom_histogram()
p = p + geom_vline(xintercept=data2plot$boot_res[1])
p = p + xlab("Bootstrap Correlation")
p
```

PLS analysis of the specific sex main effect

```{r, warning=FALSE, message=FALSE}
data2plot = read.csv(file.path(resultpath,"pls_SEX_H_brainBSR4plotting.csv"))
data2plot =  merge(data2plot,hcp_annot,by="parcel_index")


data2plot$BSR_LV1_thresh = data2plot$BSR_LV1
mask = abs(data2plot$BSR_LV1_thresh)>abs(quantile(data2plot$BSR_LV1, c(0.8)))
data2plot$BSR_LV1_thresh[!mask] = NA

data2plot = data2plot[-c(131,172),]

# show effect size throughout the entire brain
g1 = ggseg(data = data2plot, mapping=aes(fill = BSR_LV1_thresh), position = "stacked", atlas = "glasser") + theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="BSR")) +
  scale_fill_gradient2(midpoint=0, low="blue", mid="white", high="red", limits=c(-25, 100), breaks = round(seq(from=-25,to=100, by=25),2)) + guides(fill = guide_colourbar(nbin = 100))
g1
```

Plot a histogram of the correlation between H and the sex main effect across all bootstrap 10,000 resamples. The actual correlation in the real data is shown with the vertical line.

```{r, warning=FALSE, message=FALSE}
data2plot = read.csv(file.path(resultpath,"pls_SEX_H_bootres4plotting_LV1_ci95.csv"))

p = ggplot(data = data2plot, aes(x = boot_res)) + geom_histogram()
p = p + geom_vline(xintercept=data2plot$boot_res[1])
p = p + xlab("Bootstrap Correlation")
p
```


