---
title: "Overlap between ASD Genes and DHT DE genes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
easypackages::libraries("here","ggplot2","gplots","ggseg","ggsegExtra","tidyverse","matlabr")
code_dir = here("code")
source(file.path(code_dir,"genelistOverlap.R"))

genelist_dir = here("data","asd_genelists")

backgroundTotal = 13284
dht_de_genes = read.delim(file.path(genelist_dir,"dysreg_sv_limma_FDR_q0.05_resultTable.txt"))
dht_de_genes = as.character(dht_de_genes$GeneSymbol)

# ASD risk genes from Satterstrom et al., 2020, Cell
satterstrom_102_Ex = as.character(read.csv(file.path(genelist_dir,"satterstrom_2020_cell_102asdriskgenes_ExGenes.txt"),header = FALSE)$V1)
satterstrom_102_In = as.character(read.csv(file.path(genelist_dir,"satterstrom_2020_cell_102asdriskgenes_InGenes.txt"),header = FALSE)$V1)
satterstrom_102_MG = as.character(read.csv(file.path(genelist_dir,"satterstrom_2020_cell_102asdriskgenes_MicrogliaGenes.txt"),header = FALSE)$V1)
satterstrom_102_AsOPC = as.character(read.csv(file.path(genelist_dir,"satterstrom_2020_cell_102asdriskgenes_AsOPCGenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in excitatory neurons
excitatory_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_L23456degenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in inhibitory neurons
inhibitory_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_INPVSSTVIPSV2Cdegenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in Microglia
microglia_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_Microgliadegenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in Oligodendrocytes
oligodendrocyte_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_Oligodendrocytesdegenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in Astrocytes
astrocyte_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_Astrocytedegenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in OPCs
opc_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_OPCdegenes.txt"),header = FALSE)$V1)
```

## Overlap between Excitatory neurons in Velmeshev et al., 2019, Science AND any excitatory gene from Satterstrom

```{r, warning=FALSE, message=FALSE}
ExRisk_de_genes = unique(c(excitatory_de_genes,
                           satterstrom_102_Ex))
overlap_ExRisk_res = genelistOverlap(data.frame(dht_de_genes),
                                data.frame(ExRisk_de_genes),
                                backgroundTotal,
                                print_result = FALSE,
                                header = FALSE)
overlap_ExRisk_res[[1]]$OR
overlap_ExRisk_res[[1]]$hypergeo_p
sort(overlap_ExRisk_res[[1]]$overlapping_genes)
```

## Overlap between Inhibitory neurons in Velmeshev et al., 2019, Science AND any Inhibitory gene from Satterstrom

```{r, warning=FALSE, message=FALSE}
InRisk_de_genes = unique(c(inhibitory_de_genes,
                           satterstrom_102_In))
overlap_InRisk_res = genelistOverlap(data.frame(dht_de_genes),
                                data.frame(InRisk_de_genes),
                                backgroundTotal,
                                print_result = FALSE,
                                header = FALSE)
overlap_InRisk_res[[1]]$OR
overlap_InRisk_res[[1]]$hypergeo_p
sort(overlap_InRisk_res[[1]]$overlapping_genes)
```

## Overlap between Microglia cells in Velmeshev et al., 2019, Science AND any Microglia gene from Satterstrom

```{r, warning=FALSE, message=FALSE}
MGRisk_de_genes = unique(c(microglia_de_genes,
                           satterstrom_102_MG))
overlap_MGRisk_res = genelistOverlap(data.frame(dht_de_genes),
                                data.frame(MGRisk_de_genes),
                                backgroundTotal,
                                print_result = FALSE,
                                header = FALSE)
overlap_MGRisk_res[[1]]$OR
overlap_MGRisk_res[[1]]$hypergeo_p
sort(overlap_MGRisk_res[[1]]$overlapping_genes)
```

## Overlap between Astrocyte, OPC, or Oligodendrocyte cells in Velmeshev et al., 2019, Science AND any Astrocyte and OPC gene from Satterstrom

```{r, warning=FALSE, message=FALSE}
AsOPCRisk_de_genes = unique(c(oligodendrocyte_de_genes,
                              opc_de_genes,
                              astrocyte_de_genes,
                              satterstrom_102_AsOPC))
overlap_AsOPCRisk_res = genelistOverlap(data.frame(dht_de_genes),
                                data.frame(AsOPCRisk_de_genes),
                                backgroundTotal,
                                print_result = FALSE,
                                header = FALSE)
overlap_AsOPCRisk_res[[1]]$OR
overlap_AsOPCRisk_res[[1]]$hypergeo_p
sort(overlap_AsOPCRisk_res[[1]]$overlapping_genes)
```

## Make a heatmap of overlapping DHT and autism-associated excitatory genes

```{r, warning=FALSE, message=FALSE}
df = read.csv(here("data","asd_genelists","overlap_dht_de_data.csv"),row.names=1)
df = as.matrix(df)

# plot the heatmap of the data matrix
heatmap.2(df,
          Rowv=TRUE, Colv=TRUE, 
          col=bluered,
          scale="row", 
          trace="none",
          srtCol = 25)
```

## Plot percent overlap between HCP parc regions and FDR q<0.01 DHT DE and ASD E genes map

```{r, warning=FALSE, message=FALSE}
options(matlab.path = "/Applications/MATLAB_R2019b.app/bin")

RUN_MATLAB = TRUE
if (RUN_MATLAB){
  code2run = sprintf("cd %s; calculate_overlap_hcpParc_dhtasdEmap;",code_dir)
  res = run_matlab_code(code2run)
} # if (RUN_MATLAB)

dir2use = here("data","ge_maps")
df2plot = read.csv(file.path(dir2use,"hcp_fdr01_percent_overlap.csv"))

# show effect size throughout the entire brain
g1 = ggseg(.data=df2plot, 
           atlas=glasser, 
           position="stacked",
           mapping=aes(fill=percent_overlap)) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  guides(fill=guide_legend(title="% Overlap")) +
  scale_fill_gradientn(colors = colorRampPalette(c("blue","white","red"))(100),limits=c(0,100)) + 
  guides(fill = guide_colourbar(nbin = 100))
g1
```